---
title: "Visualizations"
---

<style type="text/css">
.main-container {
  max-width: 1500px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{r, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE, cache = T)
```

```{r}
library(MixSim)
library(RColorBrewer)
library(MASS)
library(expm)
library(grDevices)
library(radviz3d)
library(rgl)
library(ggplot2)
library(Rtsne)
library(Rdimtools)
library(uwot)
library(scatterplot3d) 

source("starcoord3d.R")

# color palette
load("colpal.rda")
dk.palette = c(rgb(red = 173, green = 216, blue = 230, max = 255), 
               rgb(max = 255, red = 139, green = 0, blue = 0))

```

```{r}
##
## Display the map of pairwise overlap measures of Maitra and Melnykov
## (JCGS, 2012)
##
## overlap.mat = matrix of total pairwise overlaps (necessarily symmetric)
## map.col = colormap for the mapping
## linescol = color for the lines drawing the squares
## map.range = range of the overlap map (default: minimum and maximum of lower
##               triangle of the matrix)
## lab.col = color of the labels (same as nrow(matrix) if provided)
## lab.cex = character size of the label
## map.cex = character size of the overlap values laid on the map
## legend.cex = character size of the legend text (does not work always)
##
## provides map of overlap values for each group of mixture model
##
## written Ranjan Maitra, Ames, IA 50011-1210, June 28, 2009
##
## modified Ranjan Maitra, Ames, IA 50011-1090, December 23, 2016.
## last modified Ranjan Maitra, Ames, IA 50011-1090, October 31, 2020.
##
## modification to bring in specifications for label color, labels, maps and
## legend character size; subsequently change default color of map and lines
##

overlap.map <-
    function(overlap.mat, map.col = RColorBrewer::brewer.pal(name = "PuRd", n = 9), linescol = "#1D91C0", map.range = NULL, lab = NULL, lab.col = 1, lab.cex = 5, font = 1, map.cex = 5, legend.cex = 1, scale = NULL, scale.pos = -2, legend.width = 1)
{
    oxmat <- overlap.mat
    oxmat[lower.tri(oxmat)] <- NA
    diag(oxmat) <- NA
    p <- ncol(oxmat)
    if(is.null(lab)){
      lab <- 1:p
    }
    newox <- oxmat[-p, -1]
    newox <- cbind(rbind(NA, newox), NA)[, p:1]

    layout(matrix(c(rep(1, 4*p^2), rep(2, 2*p), rep(3, 2*p), rep(4, 2*p), rep(4, 2*p)), nrow = 2*p, ncol = 2*p + 4 ))
    ##  layout(matrix(c(rep(1, p^2), rep(2, p), rep(3, p), rep(4, p)), nrow = p, ncol = p + 3))

    
    par(mar = c(0.1,0.1,0.75,0.1))
    if (is.null(map.range)) map.range <- range(newox, na.rm = T)  
    image(x = 1:p, y = 1:p, z = newox, axes = F, xlab = "", ylab = "", #col = brewer.pal(9, "GnBu"))
          col = map.col, zlim = map.range)
    text(y = 2:p, x = rep(1, p-1), labels = lab[p:2], cex = lab.cex, font = font, col = lab.col[p:2])
    text(x = 2:p, y = rep(1, (p-1)), labels = lab[1:(p-1)], cex = lab.cex, font = font, col = lab.col[1:(p-1)])
    if (!is.null(scale)){
       text(x = p-scale.pos, y = p+0.2, labels = scale, cex = map.cex)
    }
   


    for(i in 1:p) {
        for(j in i:p) {
            lines(x = c(i+0.5, i+0.5), y = c(p-j+1,p-j)+1.5, col = linescol, lwd = 0.5)
            lines(y = c(i+0.5, i+0.5), x = c(p-j+1,p-j)+1.5, col = linescol, lwd = 0.5)
        }
    }
    for(i in 2:p) text(x=1:p, y = i, labels = round(newox[,i], 2),
                       col = ifelse(newox[,i] < median(map.range), "black", "white"), cex = map.cex)

   frame()

    # savepar <- par(cex=0.75, lwd=0.25, mar = c(1, 0.5, 1, 2),
    #                 xaxs="i", yaxs="i")
    
    savepar <- par(cex=0.75, lwd=0.25, mar = c(0.5, 0.5, 0.5, 0.5),
                   xaxs="i", yaxs="i")
    # if (legend){
    plot.new()
    length.col <- length(map.col) + 1
    ra <- seq(from = map.range[1], to = map.range[2], length=length.col)
    plot.window(xlim=c(0,0.1), ylim= c(map.range[1], map.range[2]))
    rect(0, ra[-length.col], legend.width, ra[-1], col = map.col, border = NULL)
    axis(4, at = ra, labels = round(ra, digits = 2), las = 1, cex.axis = legend.cex, line = NA)
    

    rect(0, 1, legend.width, ra[length.col], col = NULL)
    frame()
    }

```


```{r, eval = F}
# Simulated Datasets
omegas <- c(0.001, 0.01, 0.05)
sim_data <- NULL
overlap_mat_sim <- NULL

seed = runif(3)*10^8
for (i in 1:length(omegas)){
  set.seed(seed[i])
  p = 50
  Q <- MixGOM(goMega = omegas[i], K = 5, p = p, sph = T, hom = T)
  overlap_mat_sim[[i]] <- overlap(Q$Pi, Q$Mu, Q$S)$OmegaMap
  #print(t(overlap(Q$Pi, Q$Mu, Q$S)$OmegaMap) + overlap(Q$Pi, Q$Mu, Q$S)$OmegaMap - diag(5))
  A <- simdataset(n = 500, Pi = Q$Pi, Mu = Q$Mu, S = Q$S)
  sn <- A$X
  class <- as.factor(A$id)
  sn.df <- data.frame(class,sn)
  sim_data[[i]] <- sn.df
}

convert_to_discrete <- function(data, discrete_p){
  n <- nrow(data)
  data[,discrete_p] <- apply(data[,discrete_p], MARGIN = 2, FUN = function(x) floor(order(x)/(n+1)*10))
  return(data)
}

for (i in 1:3){
  df <- sim_data[[i]]
  p <- ncol(df)-1
  df.mix <- convert_to_discrete(df[,-1], discrete_p = 1:round(p/2))
  sim_data[[i]] <- data.frame(class = df$class, df.mix)
}

save(overlap_mat_sim, sim_data, seed, file = "sim_data_50.rda")

# 2 groups
omegas <- c(0.001, 0.01, 0.05)
sim_data <- NULL
overlap_mat_sim <- NULL

seed = runif(3)*10^8
for (i in 1:length(omegas)){
  set.seed(seed[i])
  p = 100
  Q <- MixGOM(goMega = omegas[i], K = 2, p = p, sph = T, hom = T)
  overlap_mat_sim[[i]] <- overlap(Q$Pi, Q$Mu, Q$S)$OmegaMap
  #print(t(overlap(Q$Pi, Q$Mu, Q$S)$OmegaMap) + overlap(Q$Pi, Q$Mu, Q$S)$OmegaMap - diag(5))
  A <- simdataset(n = 500, Pi = Q$Pi, Mu = Q$Mu, S = Q$S)
  sn <- A$X
  # colnames(sn) <- paste(1:4)
  class <- as.factor(A$id)
  sn.df <- data.frame(class,sn)
  sim_data[[i]] <- sn.df
}

convert_to_discrete <- function(data, discrete_p = 1:50){
  n <- nrow(data)
  data[,discrete_p] <- apply(data[,discrete_p], MARGIN = 2, FUN = function(x) floor(order(x)/(n+1)*10))
  return(data)
}

for (i in 1:3){
  df <- sim_data[[i]]
  df.mix <- convert_to_discrete(df[,-1])
  sim_data[[i]] <- data.frame(class = df$class, df.mix)
}

save(overlap_mat_sim, sim_data, seed, file = "sim_data_100_2g.rda")

# 3 groups
omegas <- c(0.001, 0.01, 0.05)
sim_data <- NULL
overlap_mat_sim <- NULL

seed = runif(3)*10^8
for (i in 1:length(omegas)){
  set.seed(seed[i])
  p = 100
  Q <- MixGOM(goMega = omegas[i], K = 3, p = p, sph = T, hom = T)
  overlap_mat_sim[[i]] <- overlap(Q$Pi, Q$Mu, Q$S)$OmegaMap
  #print(t(overlap(Q$Pi, Q$Mu, Q$S)$OmegaMap) + overlap(Q$Pi, Q$Mu, Q$S)$OmegaMap - diag(5))
  A <- simdataset(n = 500, Pi = Q$Pi, Mu = Q$Mu, S = Q$S)
  sn <- A$X
  # colnames(sn) <- paste(1:4)
  class <- as.factor(A$id)
  sn.df <- data.frame(class,sn)
  sim_data[[i]] <- sn.df
}

convert_to_discrete <- function(data, discrete_p = 1:50){
  n <- nrow(data)
  data[,discrete_p] <- apply(data[,discrete_p], MARGIN = 2, FUN = function(x) floor(order(x)/(n+1)*10))
  return(data)
}

for (i in 1:3){
  df <- sim_data[[i]]
  df.mix <- convert_to_discrete(df[,-1])
  sim_data[[i]] <- data.frame(class = df$class, df.mix)
}

save(overlap_mat_sim, sim_data, seed, file = "sim_data_100_3g.rda")

# 4 groups
omegas <- c(0.001, 0.01, 0.05)
sim_data <- NULL
overlap_mat_sim <- NULL

seed = runif(3)*10^8
for (i in 1:length(omegas)){
  set.seed(seed[i])
  p = 100
  Q <- MixGOM(goMega = omegas[i], K = 4, p = p, sph = T, hom = T)
  overlap_mat_sim[[i]] <- overlap(Q$Pi, Q$Mu, Q$S)$OmegaMap
  #print(t(overlap(Q$Pi, Q$Mu, Q$S)$OmegaMap) + overlap(Q$Pi, Q$Mu, Q$S)$OmegaMap - diag(5))
  A <- simdataset(n = 500, Pi = Q$Pi, Mu = Q$Mu, S = Q$S)
  sn <- A$X
  # colnames(sn) <- paste(1:4)
  class <- as.factor(A$id)
  sn.df <- data.frame(class,sn)
  sim_data[[i]] <- sn.df
}

convert_to_discrete <- function(data, discrete_p = 1:50){
  n <- nrow(data)
  data[,discrete_p] <- apply(data[,discrete_p], MARGIN = 2, FUN = function(x) floor(order(x)/(n+1)*10))
  return(data)
}

for (i in 1:3){
  df <- sim_data[[i]]
  df.mix <- convert_to_discrete(df[,-1])
  sim_data[[i]] <- data.frame(class = df$class, df.mix)
}

save(overlap_mat_sim, sim_data, seed, file = "sim_data_100_4g.rda")
```


## Fig. S1: Overlap Map, starcoords3D and UMAP of Simulated 100D Mixed Datasets with 5 Groups

(a) [Overlap Map](sim_overlap.html)
(b) [starcoords3D](sim_starcoords3d.html)
(c) [UMAP](sim_umap.html)

## Fig. S2: starcoords3D and 2D/3D Scatter Plots of Simulated 100D Mixed Datasets with 2 Groups

(a) [starcoords3D](sim_starcoords3d_2g.html)
(b) [2D Scatter Plot](sim_scatter2d_2g.html)
(c) [3D Scatter Plot](sim_scatter3d_2g.html)

## Fig. S3: Overlap Map, starcoords3D and 2D/3D Scatter Plots of Simulated 100D Mixed Datasets with 3 Groups

(a) [Overlap Map](sim_overlap_3g.html)
(b) [starcoords3D](sim_starcoords3d_3g.html)
(c) [2D Scatter Plot](sim_scatter2d_3g.html)
(d) [3D Scatter Plot](sim_scatter3d_3g.html)

## Fig. S4: Overlap Map, starcoords3D and 2D/3D Scatter Plots of Simulated 100D Mixed Datasets with 4 Groups

(a) [Overlap Map](sim_overlap_4g.html)
(b) [starcoords3D](sim_starcoords3d_4g.html)
(c) [2D Scatter Plot](sim_scatter2d_4g.html)
(d) [3D Scatter Plot](sim_scatter3d_4g.html)


# Real Data Applications

## Fig. S6: Senators Dataset
[t-SNE, UMAP, OLDA, ULDA, 2D&3D Scatter Plots, Starcoords3D](senator_mixedviz.html)

## Fig. S7: Adult Autism Spectrum Disorder (ASD) Dataset
[t-SNE, UMAP, OLDA, ULDA, 2D&3D Scatter Plots, Starcoords3D](autisim_mixedviz.html)

## Fig. S8: SPECT Heart Dataset
[t-SNE, UMAP, OLDA, ULDA, 2D&3D Scatter Plots, Starcoords3D](spect_mixedviz.html)

## Fig. S9: Indic Scripts Dataset
[t-SNE, UMAP, OLDA, ULDA, Starcoords3D](script_mixedviz.html)

## Fig. S10: Human Tissue RNA-seq Dataset
[UMAP, Starcoords3D](gtex_mixedviz.html)

## Fig. S11: HIV DNA-seq Dataset
[UMAP, Starcoords3D](hiv_mixedviz.html)

## Fig. S12: Mouse Brain Single-cell RNA-seq Dataset
[UMAP, Starcoords3D](braincell_mixedviz.html)


